---
title: " "
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: false
    #css: "./www/bootstrap.css"
    
---

# Introduction

*The **Social Protection Tool** provides simulations of the cost and impact of tax-financed cash transfer programmes across a number of countries in Asia and the Pacific. It can be used to make initial estimates of cost projections for different social protection programme schemes and their impact on coverage, household purchasing power, poverty and inequality. The tool also illustrates programme affordability by comparing estimated budget against alternative financing streams.* 

*Users can develop different policy options by simulating the introduction of social security programmes. The tool includes two types of schemes: inclusive programmes organised around the lifecycle (children, older persons, and persons with disabilities); and, means-tested household programmes. Users can set parameters related to eligibility criteria, desired levels of programme coverage, and transfer values. Users may also modify parameters related to the expected real annual growth rate of the economy, and administrative programme costs.*

Once parameters of the selected programmes have been defined and set, the tool provides the results to the users through interactive and easy-to-understand tables and graphs. The tool also allows the user to download a pdf report with all the key results and charts.

The simulations will allow users to observe what would have happened, in terms of household welfare, if such programmes had been in place in the year of the household survey. The simulated impacts focus on showing changes in key indicators related to programme coverage and the estimated impacts on poverty and inequality. The indicators are:


* **Programme coverage:** the estimated percentage of the population eligible for the simulated programme(s), in total and disaggregated by deciles of household income (or expenditure), location (urban-rural) and household living arrangement type.
* **Impact on purchasing power:** the estimated increase in households’ consumption expenditure as a result of the social protection programme(s), disaggregated by deciles of household income (or expenditure), location (urban-rural) and household living arrangement type.
* **Impact on poverty:** the estimated reduction in levels of poverty among recipients of the programme(s) and among the general population as measured by the Foster-Greer-Thorbecke (FGT) class of poverty measures (headcount ratio and poverty gap), using different poverty lines. These poverty measures are compared in the context of no schemes and with any existing country schemes.
* **Impact on inequality:** the estimated reduction in inequality as measured by different metrics, such as the Gini index, the Quintile ratio and the Palma ratio.

In countries with no or limited social protection coverage, it is often argued that social protection is not affordable and that governments have insufficient fiscal space to introduce new programmes. The tool, therefore, also computes the relative cost of simulated programme(s) and provides different illustrative options available to expand fiscal space for social protection. The tool provides the following cost and affordability indicators:

* **Projected annual cost of the simulated programme(s):** the estimated projected annual costs up to 2030 of the programme(s), provided both in absolute values and expressed as a percentage of gross domestic product (GDP).
* **Required investment as a percentage of government revenue:** the estimated projected annual costs expressed as a percentage of today’s government revenue.
* **A comparison of the estimated cost of the simulated programme(s) with public spending on other sectors:** the estimated total cost of the programme relative to other functions of government spending, such as defence, health and education.
* **A comparison of the estimated cost of the simulated programme(s) with government revenue generated by income and corporate taxes: ** the estimated total cost of the programme relative to how much tax revenue would have to increase in order to invest in the desired reforms..

All the simulations built into the tool are computed using R, an open source statistical software package. The online tool was developed using Shiny which is an R package developed by RStudio. Below is a detailed description of the methods and assumptions used in the tool to simulate impacts and project costs and affordability indicators.

# Methods and Assumptions

## Simulating impacts

The simulations in the tool answer *“what if”* questions in a static and backward-looking manner. Using nationally representative household surveys, the simulations provide hypothetical calculations of how the introduction of tax-financed cash transfer programmes impact the welfare of household units (i.e., those directly and indirectly benefiting from the programmes). The simulations attempt to construct a hypothetical scenario of what would have happened to households if such programmes had been in place in the year of the household survey. By looking at the household unit, distributional effects of the simulated impact estimates can also be analysed. For instance, the impact estimates are also presented by age groups and welfare quantiles.

Behind these hypothetical calculations are a number of assumptions. The main assumption in the microsimulation model is that households expend 100 per cent of all additional income from the cash transfers. That is, the model does not incorporate other possible behavioural responses to these potential changes in household income. For example, as a result, in the simulations, households do not save any portion of the transfers received.

Baseline and counterfactual scenarios are established in order to infer the absolute effects of a policy change. The model itself is a linear approximation model as outlined by Figari, Paulus and Sutherland (2015)^[Figari, F., Paulus, A., & Sutherland, H. (2015). Microsimulation and policy analysis. In Handbook of income distribution (Vol. 2, pp. 2141-2221). Elsevier.] that decomposes household expenditure to isolate the effect of a cash transfer benefit to the household, conditional on a set of household socio-demographic characteristics, as dictated by the Tool user.

The simulation model uses per capita consumption expenditure as the measure for household welfare and looks at impacts across four different sets of outcomes: programme coverage, household consumption expenditure levels, poverty and inequality outcomes. Five basic steps where generally taken to estimate the simulated impacts in terms of these outcomes. The steps are described below.

### Steps {.tabset}

#### Step 1
***Deduct existing tax-financed cash transfer programmes from the households’ current expenditure levels***

When possible, that is, when data permit, transfer values of existing non-contributory programmes that are similar to the proposed programmes are deducted from current expenditure levels.

Below is a list of all programmes that were possible to deduct from households’ current expenditure levels. This is also highlighted in the tool. The user is informed of any existing non-contributory cash transfer programmes once a scheme has been selected. An info box appears whenever a country has a programme that is similar in nature.

##### Existing programmes {.tabset}

###### Bangladesh

* *Old Age Allowance* - Old age benefit

###### Georgia

* *Old Age Pension (OAP)* - Old age benefit
* *Targeted Social Assistance (TSA)* - Houseshold benefit

###### India

* *Indira Gandhi National Old Age Pension Scheme (IGNOAPS)* - Old age benefit
* *Indira Gandhi National Disability Pension Scheme (IGNDPS)* - Disability benefit

###### Indonesia

* *Program Keluarga Harapan (PKH)* - Houseshold benefit

###### Kyrgyzstan

* *Basic Old Age Pension* - Old age benefit
* *Disability Allowance* - Disability benefit

###### Maldives

* *Single Parent Allowance* - Child benefit
* *Disability Allowance programme* - Disability benefit
* *Old Age Basic Pension* - Old age benefit

###### Mongolia

* *Child Money Programme* - Child benefit

###### Nepal 

* *Disability Allowance* – Disability benefit
* *Senior Citizen's Allowance* – Old age benefit

###### Pakistan

* *Benazir Income Support Programme (BISP)* - Houseshold benefit

###### Philippines

* *Pantawid Pamilyang Pilipino Program (4Ps)* - Houseshold benefit

###### Sri Lanka

* *Samurdhi* - Houseshold benefit

###### Thailand
* *Social Pension* – Old age benefit

###### Viet Nam

* *Old Age Pension* - Pension-tested for older people 80+ years and targeted older people 60 - 79 years

#### Step 2

***Identify beneficiaries of the programmes***

From the parameters set by the user for each programme, age-eligible individuals and households are identified in the dataset. When the programme is not universal (i.e., coverage is not 100 per cent) it becomes means-tested. However, in order to mimic current practice, selection of age-eligible individuals and households in means-tested programmes are identified using the proxy means test (PMT) approach. More detail on the PMT used in each country is described below.

Because not all datasets include disability questionnaires, persons with disabilities are identified using the following approach:

```{r echo=FALSE}
library(DT)
countrylist <-
  c("Bangladesh","Georgia",
    "India",
    "Indonesia",
    "Kyrgyzstan",
    "Maldives",
    "Mongolia",
    "Nepal",
    "Pakistan",
    "Philippines",
    "Sri Lanka",
    "Thaliand",
    "Viet Nam")
data <- read.csv("./documentation/disabAbout.csv")
names(data) <- c("Disability", countrylist)
DT::datatable(
      data,
      rownames = FALSE,
      colnames = c("Disability", countrylist),
      options = list(
        autoWidth = TRUE,
        columnDefs = list(
          list(
            className = 'dt-center',
            targets = c(1:7),
            width = '200px'
          ),
          list(
            className = 'dt-left',
            targets = c(0),
            width = '500px'
          )
        ),
        dom = 't',
        ordering=F
      ),
      escape = FALSE
    )
```
**_Note_**: a) in Mongolia, standard ‘Yes/No’ question of whether the respondent has a disability; b) In India, eight questions on activities of daily living were applied to all those aged 8 and over. A person with disability is then defined as someone who answered ‘unable to do it’ to at least one activity. For children aged 7 years and under, the model assumes a disability prevalence of 1 per cent, following international estimates.

#### Step 3

***Calculate hypothetical household consumption expenditure***

Once potential recipients and their households have been identified, the specified monthly transfer values for the selected programme(s) are distributed to the recipients. All transfer values are aggregated at the household level before constructing household per capita monthly transfer values. Then, new post-reforms (ex-post) values of household per capita expenditure are estimated by first deducting existing cash transfer programmes (as described in Step 1) and adding the per capita transfer values simulated from each of the selected programmes.

#### Step 4 {.tabset}

***Simulate changes in welfare-based outcomes***

##### Coverage

Once the user selects the parameters of the programmes and beneficiaries are identified; coverage for a given programme is then estimated as the share of the eligible population living in households with at least one recipient. Coverage is calculated by household expenditure deciles, by urban/rural location and household living arrangement types.

##### Household consumption

The increase in household purchasing power is estimated by calculating the average of per capita transfer value as a percentage of household per capita consumption expenditure.^[When possible, household consumption is calculated and adjusted for regional prices and then the per capita value is calculated by dividing this figure by household size.] Transfers that the household is receiving from any similar existing programmes are taken in to account, such that the increase in household purchasing power is net of any existing benefits. Estimates are provided by expenditure deciles, urban/rural localtion and household living arrangement types, for both the total household population and beneficiary households.


##### Poverty outcomes

Two poverty outcome measures are considered. Poverty headcount of the general population and poverty gap of the general population. For both of these outcomes household consumption expenditure is used as the measure of welfare and four absolute poverty lines are used: the national poverty line and three international poverty lines. The international poverty lines are: PPP\$ 1.9 per person per day, PPP\$ 3.2 per person per day, and PPP\$ 5.5 per person per day. The poverty headcount index and poverty gap are estimated for both before and after the proposed reforms by the user.

##### Inequality outcomes

Three generally accepted inequality metrics are used in the tool. All of them use household consumption expenditure as the measure of welfare. The metrics are:

* Gini index (ratio of the area between the Lorenz curve which depicts the percentage of income owned by x per cent of the population. and the 45-degree line which represents perfect equality, and the area beneath the 45-degree line)
* Quintile ratio (or 20:20 ratio i.e. the average national income of the top 20 per cent to the bottom 20 per cent)
* Palma ratio (or 10:40 ratio which is the national income share of the top 10 per cent to the bottom 40 per cent)

#### Step 5 {.tabset}

***Projecting costs and affordability indicators***

In order to highlight the sustainability of the proposed reforms in terms of its costs and affordability, the tool also projects the cost going forward and provides different illustrative options to expand fiscal space for social protection. Below is a description of how programme costs are projected and how the affordability measures are calculated. Some of the affordability indicators are derived from Ortiz, Cummins and Karunanethy 2017 working paper Fiscal space for social protection and the SDGs: Options to expand social investments in 187 countries.

***Projected annual cost of the simulated programme(s)***

The costs in real values (in 2020 prices) are projected using UN-DESA’s Population Prospects 2017 revision data and the selected parameters of the proposed programme(s): eligibility criteria, monthly transfer values and administrative cost. The UN-DESA’s data provide for each of the countries in the tool, the projected total number of people in each year going forward by single age groups. The estimated projected annual costs in the tool are for the years 2020 to 2030.

While projecting annual costs of the child and old age benefit programmes is straightforward in the tool, it is not for the means-tested household programme and the disability benefit programme. For the child and old age benefit programmes, the total annual costs in 2020 prices is simply the product of the following factors: projected total number of individuals that fall under the age-eligibility criterion 〖Pop〗_t^j, coverage criterion, Criterion_j, as the proportion of the total number of individuals i  that fall under the age-eligibility criteria in a given time period, annualised transfer value in 2020 prices, and one plus the administrative cost as a proportion of the total transfer cost.^[Costs are calculated in the same way from 2020 to 2030 and transfer values and, by extension, administrative costs are constant overtime]

However, because the population projection data do not provide the projected total number of persons with disabilities by single age groups, nor the projected total number of households, additional assumptions are required to project the annual costs of the means-tested household programme and the disability benefit programme.

To project the total number of persons with disabilities by single age groups, the tool assumes constant disability prevalence rates by single age groups. The rates are assumed to be the same as those in the household surveys or constructed into the survey datasets as explained in Step 2 of the algorithm used to simulate impacts. Once, the total number of households and persons with disabilities by single age groups have been projected into the future, then the projection of the annual costs of the means-tested household programme and the disability benefit programme follows the same approach used to project the annual costs of the child and old age benefit programmes.

In the means-tested household programme, the total number of households in a country is projected by dividing its total projected population in a given year by the projected average household size for that same year.

Projecting population of households is complex and requires information on trends in household structure. However, there is a world trend in declining household size in low and middle countries which we attempt to factor in. A simplified approach is adopted where the projected number of households for a given year is calculated as the number of households in 2020 times the share of adults (18 – 69 years) in that year divided by the share of adults (18 – 69 years) in 2020, such that the ratio the number of households in a given year to the number of households in the base year is the same as the ratio of the number of adults in the given year to the number of adults in the base year.

The projected annual costs are also expressed as a percentage of gross domestic product (GDP). These are estimated by dividing the projected annual costs of each of the programmes in 2020 prices by the projected GDP of the country in 2020 prices. To project the GDP in real values forward, the tool uses the 2020 GDP estimates in the IMF’s World Economic Outlook (WEO) database and the average annual real growth for the period 2020 to 2023 also in the WEO.

***Required investment as a percentage of government revenue***

The first affordability indicator shows the total cost of the proposed reforms as a percentage of government revenue. This is then projected forward by assuming government revenue to be a constant share of GDP. Estimates of government revenue as a percentage of GDP are from the IMF’s World Economic Outlook database.

***A comparison of the estimated cost of the simulated programme(s) with public spending on other sectors***

The second indicator compares the estimated total cost of the proposed reforms in 2020 with government estimated spending on key functions in 2020. The key functional government expenses follow those that are presented in the IMF’s Government Finance Statistics (GFS) database. However, most countries did not have available data on government functional expenses in the GFS database. This was the case for India, Indonesia, Pakistan, Philippines and Viet Nam. For these countries, government spending in key functions were retrieved from the latest Annual Budgets available for each of these countries.

***A comparison of the estimated cost of the simulated programme(s) with government revenue generated by income and corporate taxes***

A third affordability indicator looks at how much the relevant government would be required to increase revenue from income and corporate taxes by. This is demonstrated by showing the total cost of the proposed reforms in 2020 as a percentage of government’s estimated income and corporate tax revenue in 2020. Estimates of government income and corporate tax revenues as a percentage of GDP were extracted from country specific Annual Budgets.


### Proxy means test  {.tabset}

#### PMT

As discussed in Step 2 of the simulation algorithm, proxy means test (PMT) are used to rank households before selection when universal coverage is not considered. Following standard practice, PMTs in this tool are developed using regression models of household welfare on a select number of welfare related covariates (proxies). In addition to goodness-of-fit statistics of the regression models, separate estimates of regression coefficients for each country are presented in the following tab.

#### Regression outputs


```{r echo=FALSE}
library(DT)
countrylist <-
  c("Bangladesh","Georgia",
    "India",
    "Indonesia",
    "Kyrgyzstan",
    "Maldives",
    "Mongolia",
    "Nepal",
    "Pakistan",
    "Philippines",
    "Sri Lanka",
    "Thaliand",
    "Viet Nam")
data <- read.csv("./documentation/resultsAbout.csv")
DT::datatable(
      data,
      rownames = FALSE,
      colnames = c("VARIABLES", countrylist),
      options = list(
        autoWidth = TRUE,
        columnDefs = list(
          list(
            className = 'dt-right',
            targets = c(1:7),
            width = '200px'
          ),
          list(
            className = 'dt-left',
            targets = c(0),
            width = '500px'
          )
        ),
        ordering=F
      ),
      escape = FALSE
    )
```

(i) Whether household owns a flushing toilet
(ii) Type of toilet system 1 - Sewage/central system; 2 - Latrine (any type); 3 - Septic tank/double septic tank; - 4 - No toilet facilities/bush/field; 5 - Other (bucket, soak pit, etc)
(iii) Dwelling type: 1- Ger; 2 - Apartment; 3 - Detached house; 4 - Separate housing; 5- Student dormitory; 6 - Public dwelling for employees; 7 - Other public dwelling; 8 - Non-living quarters; 9 - Other
(iv) Type of work: 1 - Self employed; 2- Employer assisted by unpaid workers; 3 - Employer assisted by paid workers; 4 - Employee; 5 - Freelance; 6 - Unpaid family worker 
(v) Pakistan assets are calculated in real terms 


# Data sources {.tabset}

## Economic data

* International Monetary Fund, World Economic Outlook Database, October 2019

## Population projection data

  * UN-DESA Population Prospects 2019 revision

## Household surveys

  * Bangladesh: HIES (2016)	
  * Georgia: WMS (2015)	
  * India: IHDS (2012)	
  * Indonesia: SUSENAS (2017)
  * Kyrgyzstan: KHS (2018)
  * Maldives: HIES (2016)	
  * Mongolia:	HSES (2016)
  * Philippines: APIS (2014)
  * Pakistan:	HIES (2015-2016)
  * Nepal: NLSS (2010/2011)	
  * Sri Lanka:	HIES (2016)
  * Thailand: SES (2018)	
  * Viet Nam: VHLSS (2016)	


## Other data sources

  * IMF - Government Finance Statistics (GFS) data, 2019
  * Estimates on functional expenses for India, Indonesia, Pakistan, Philippines and Viet Nam are from their latest Annual Budgets.

# Limitations

This tool has a few limitations. The limitations derive from the assumptions and datasets used to run the tool. Below is a description of the main limitations.
  
**Limitations with the assumptions**

The assumptions utilized vary in terms of how restrictive they are , and, ultimately affect how accurately the tool can estimate impact, costs and affordability. Thus, the user should see the model as a tool that can guide initial discussions in reforming social protection programmes in these countries, rather than a tool that provides precise estimates of impact, costs and affordability.

A key limitation of the tool is how persons with disabilities are identified in the simulations when household surveys do not provide measures of disability. The model takes on a very simplistic approach as it only looks at one dimension – age – to assign persons with disabilities. This potentially has great repercussions on the simulated impact results that either look directly or indirectly at the welfare of the households, for example simulated coverage by welfare deciles. However, because persons with disabilities usually represent a very small share of the total population, especially of those under 60 years of age, these limitations are less likely to impact the overall estimates of the reforms.

It should also be highlighted that the affordability indicators in the tool do not take into account existing government expenditure on social protection programmes that are similar to those in the tool. Therefore, some of these indicators are potentially underestimating governments’ real fiscal space capacity to invest in social protection programmes.

**Limitations with the datasets**

**Indonesia**: Because of usability of the tool and the time it takes to carry out the simulations using the full sample in the 2017 SUSENAS, the simulations for Indonesia are only carried out in 10 per cent of the sample in the SUSENAS dataset.

**Mongolia**: For the most part, the household surveys used for the simulations included household expenditure aggregates that were constructed by the dataset proprietor. This meant that it was possible to replicate welfare and poverty levels outlined in their respective official reports. However, this was not the case for Mongolia, where household expenditure aggregates had to be computed following the methodology outlined in Annex B of the ‘2016 Poverty Profile’ of the Household Socioeconomic Survey (HSES) 2016 final report. The household consumption variable consisted of consumption from food, non-food, durable goods and housing rents, which were aggregated at the household level to construct household monthly consumption expenditure per capita. The methodology as outlined in Annex B was followed as closely as possible, however, the constructed variable is 10 per cent higher than the value presented in the Poverty Profile report. Therefore, in order to match official poverty estimates, a new poverty line was constructed using the reported poverty headcount rate and regional price adjustment factor, as outlined in the report. When looking at the key consumption groups, estimated housing rent and consumption stream derived from durable goods were especially difficult to replicate since both of these consumption subcategories had been derived from linear regressions. The implication of this is that it was not possible to use the National Poverty Line used in the Poverty Profile report of MNT 145,146 giving a headcount poverty rate of 29.6%. For the purposes of this tool, it was sufficient to impose a poverty rate of 29.6% using the computed household consumption variable and a poverty line of MNT 136,815.

**Georgia**: The 2015 Welfare Monitoring Survey poverty estimates are provided using per adult equivalent household consumption expenditure. This tool, however, uses per capita household consumption expenditure as the measure of welfare. Therefore, in order to maintain the same reported poverty levels and use per capita measures of consumption expenditure, a new poverty line had to be created

# Acronyms

```{r echo=FALSE}
library(DT)
data <- read.csv("./documentation/acronymsTable.csv")
DT::datatable(
      data,
      rownames = FALSE,
      options = list(
        autoWidth = TRUE,
        columnDefs = list(
          list(
            className = 'dt-left',
            targets = c(0),
            width = '200px'
          ),
          list(
            className = 'dt-left',
            targets = c(1),
            width = '500px'
          )
        ),
        dom = 'ft',
        pageLength = 100
      ),
      escape = FALSE
    )
```

# Notes
